# The Area Dataset

```{r}
library(magrittr)
library(tidyverse)
load("../output/essential_aera_raion_demo_macro.RData")
```

The `area` dataset is the largest, containing `r ncol(area)` columns. All the variables containing `count` in the names are, well, counts. The variables starting with `ID_` are actually categorical and not numeric. There are `r sum(grepl('^ID_', names(area)))` such columns. Intuitively, it seems more relevant that there is a bus/train/metro station nearby, rather than the actual number of the line. We drop these columns since they contain quite a lot of values (see below)

```{r}
area %>% select(starts_with("ID_")) %>% map(unique)
area %<>% select(-starts_with("ID_"))
```

There are a few character columns. They seem to be mostly binary variables

```{r}
area_char <- area %>% map_lgl(is.character) %>% which()
area_char
```

Let's verify that they are actually binary

```{r}
area %>% select(area_char) %>% map_int(~length(unique(.x)))
```
Yes, they are all binary. We convert them to dichotomous variables.

```{r}
foo <- function(x) {ifelse(x == 'yes', 1, 0)}
area %<>% mutate(
   culture_objects_top_25 = foo(culture_objects_top_25),
   water_1line = foo(water_1line),
   big_road1_1line = foo(big_road1_1line),
   railroad_1line = foo(railroad_1line)
)
any(map_lgl(area, is.character))
```

## Selecting the area variables with XGBoost

If we run an xgboost model on the `area` dataset, and try to predict the prices on the pseudo-test set shown below, we find that the best number of rounds is around 720. 
```{r, eval=TRUE, cache=TRUE}
library(xgboost)
X <- as.matrix(area)
y <- y_train

params <- list(
  eta = 0.1, 
  gamma = 1,
  max_depth = 5,
  min_child_weight = 1,
  subsample = 0.9,
  colsample_bytree = 0.9,
  objective = "reg:linear")

dTr <- xgb.DMatrix(X, label = y)

set.seed(123)
cv <- xgb.cv(params = params, data = dTr, nrounds = 500, nfold = 5)
tr <- xgb.train(params = params, data = dTr, 
                nrounds = which.min(cv$evaluation_log$test_rmse_mean))
imp <- xgb.importance(colnames(X), model = tr)
xgb.plot.importance(imp)
to_keep <- filter(imp, Importance > 0.01)
area <- select(area, one_of(to_keep$Feature))
```

To start with, we can retain only the variables with an importance above 0.01. This has left `r ncol(area)` features in the area dataset.

```{r}
save(area, file = "../output/04-area_dataset.RData")
```

