# The Macro Dataset

```{r}
library(magrittr)
library(tidyverse)
load("../output/essential_aera_raion_demo_macro.RData")
macro <- inner_join(prices, macro)
```

We remove the id variable and add some time-related variables, namely year, month and week.

```{r}
macro_timestamp <- macro$timestamp
macro %<>% select(-id) %>%
  mutate(year = lubridate::year(timestamp),
         month = lubridate::month(timestamp),
         week = lubridate::week(timestamp),
         price = log(1 + price_doc)) %>%
  select(-timestamp)
```

## XGBoost model

```{r}
library(xgboost)
y <- macro$price
X <- macro %>% select(-c(price_doc, price)) %>% as.matrix()

params <- list(
  eta = 0.1, 
  gamma = 1,
  max_depth = 5,
  min_child_weight = 1,
  subsample = 0.9,
  colsample_bytree = 0.9,
  objective = "reg:linear")

dTr <- xgb.DMatrix(X, label = y)

set.seed(123)
cv <- xgb.cv(params = params, data = dTr, nrounds = 300, nfold = 5)
tr <- xgb.train(
  params = params, 
  data = dTr, 
  nrounds = which.min(cv$evaluation_log$test_rmse_mean)
)
imp <- xgb.importance(colnames(X), model = tr)
xgb.plot.importance(imp)
xgb_to_keep <- filter(imp, Importance > 0.01)
xgb_to_keep <- xgb_to_keep$Feature
```

## Glmnet

```{r}
library(glmnet)
library(caret)
bc <- preProcess(X, method = "BoxCox")
X <- predict(bc, X)
idx <- complete.cases(X)
X <- X[idx, ]
y <- y[idx]
cv <- cv.glmnet(X, y, type.measure = "mse")
plot(cv)
```

It seems that there are just 10 variables or so that are somewhat relevant.

```{r}
fit <- glmnet(X, y, lambda = cv$lambda.min)
tmp <- coef(fit)
glm_to_keep <- rownames(tmp)[tmp[, 1] != 0][-1] # Intercept
```

### Final selection

We keep the predictors selected by both methods

```{r, eval=FALSE}
to_keep <- unique(c(xgb_to_keep, glm_to_keep))
macro %<>% select(one_of(to_keep))
macro$timestamp <- macro_timestamp
save(macro, file = "../output/05-macro_dataset.RData")
```

### Alternative Approach

We select the variables recommended by Roberto Ruiz in [this kernel](https://www.kaggle.com/robertoruiz/dealing-with-multicollinearity/notebook/notebook) (look at the bottom of the post).

```{r}
vif_to_keep <- c(
  "year",
  "balance_trade",
  "balance_trade_growth",
  "eurrub",
  "average_provision_of_build_contract",
  "micex_rgbi_tr",
  # "micex_cbi_tr",
  "deposits_rate",
  "mortgage_value",
  "mortgage_rate",
  "income_per_cap",
  # "rent_price_4.room_bus",
  # "museum_visitis_per_100_cap",
  "apartment_build"
)

to_keep <- union(xgb_to_keep, vif_to_keep)
macro %<>% select(one_of(to_keep))
macro$timestamp <- macro_timestamp
save(macro, file = "../output/05-macro_dataset.RData")
```
