# The Demographic Dataset

We want to know whether the demographic information is constant for a given raion for a given year. We expect the demographic information to be raion and time specific.

```{r}
library(magrittr)
library(tidyverse)
load("../output/essential_aera_raion_demo_macro.RData")
demographic <- bind_cols(prices[, -2], demographic)
demographic$year <- lubridate::year(demographic$timestamp)
demographic$sub_area <- essential$sub_area
# demographic$timestamp <- NULL
```

We can check whether there is one record for each combination of year and `sub_area` or if there are more.

```{r}
tmp <- select(demographic, -c(id, timestamp, price_doc)) %>%
  arrange(sub_area, year) %>%
  distinct()
tmp %>% select(year, sub_area) %>% duplicated() %>% any()
```

As for the raion, the correspondence is one-to-one. It does not make sense to check the association of these variables with each single transaction, but rather to compare it with the average prices per year and per raion.

```{r}
tmp <- select(demographic, -c(id, timestamp))
tmp$full_sq <- essential$full_sq
yearly_price_demo <- tmp %>% 
  group_by(year, sub_area) %>%
  mutate(mean_price_sqm = mean(price_doc / full_sq)) %>%
  select(-c(price_doc, full_sq)) %>%
  ungroup() %>%
  distinct()
```

We can see whether there is any association between the mean log price per square meter and the raion parameters.

### GLMNET model


```{r}
library(caret)
y <- log(1 + yearly_price_demo$mean_price_sqm)
X <- yearly_price_demo %>% select(-c(mean_price_sqm, sub_area))
X <- as.matrix(X)
idx <- match("year", colnames(X))
bc <- preProcess(X[, -idx], method = "BoxCox")
X[, -idx] <- predict(bc, X[, -idx])
```

We can now fit a GLMNET model, but we need to remove the missing values first.

```{r glmnet_rion}
library(glmnet)
idx <- complete.cases(X) & complete.cases(y)
X <- X[idx, ]
y <- y[idx]
set.seed(666)
cv <- cv.glmnet(X, y)
fit <- glmnet(X, y, lambda = cv$lambda.min)
cf <- coef(fit)
glm_to_keep <- rownames(cf)[cf[, 1] != 0][-1]
```

### XGBoost Model

```{r, results=FALSE, message=FALSE}
library(xgboost)

params <- list(
  eta = 0.1, 
  gamma = 1,
  max_depth = 5,
  min_child_weight = 1,
  subsample = 0.9,
  colsample_bytree = 0.9,
  objective = "reg:linear")

dTr <- xgb.DMatrix(X, label = y)

set.seed(666)
cv <- xgb.cv(params = params, data = dTr, nrounds = 300, nfold = 5)
tr <- xgb.train(
  params = params, 
  data = dTr, 
  nrounds = which.min(cv$evaluation_log$test_rmse_mean)
)
imp <- xgb.importance(colnames(X), model = tr)
xgb.plot.importance(imp)
xgb_to_keep <- filter(imp, Importance > 0.01)
xgb_to_keep <- xgb_to_keep$Feature
```

As in the previous case, we retain the union of the variables

```{r}
to_keep <- union(glm_to_keep, xgb_to_keep)
to_keep <- gsub("yes$", "", to_keep)
stopifnot(all(to_keep %in% colnames(yearly_price_demo)))
yearly_price_demo %<>% select(sub_area, one_of(to_keep))
prices %<>% mutate(
  year = lubridate::year(timestamp),
  sub_area = demographic$sub_area
)
demographic <- inner_join(prices, yearly_price_demo,
                            by = c("year", "sub_area"))
save(demographic, file = "../output/07-demographics_dataset.Rmd")
```

