# Final Dataset

We load the datasets we have created so far and put together whatever has passed the filtering. There are still a number of operations to take care of:

1. Creating the dummy variables.
2. Box-Cox tranformation of the predictors.
3. Imputation

```{r}
library(magrittr)
library(tidyverse)
load("../output/03-essential_dataset.Rmd")
load("../output/04-area_dataset.RData")
load("../output/05-macro_dataset.RData")
load("../output/06-raion_dataset.Rmd")
load("../output/07-demographics_dataset.Rmd")
```

We first make sure that each subset has the same number of rows. We then remove the columns that appear in each subset.

```{r}
stopifnot(all.equal(
  nrow(essential),
  nrow(area),
  nrow(raion),
  nrow(demographic),
  nrow(macro),
  nrow(prices) 
))

prices$sub_area <- essential$sub_area
to_remove <- c('id', 'price_doc', 'year', 'timestamp', 'sub_area')
raion %<>% select(-one_of(to_remove))
demographic %<>% select(-one_of(to_remove))
macro %<>% select(-c(year, timestamp))
essential %<>% select(-sub_area)
```

We can now arrange all the datasets in one single tibble.

```{r make_tibble}
train_set <- bind_cols(
  essential,
  area,
  raion,
  demographic,
  macro
)
train_set <- bind_cols(prices[, 1:2], train_set)
if(any(duplicated(names(train_set))))
  stop("Duplicated columns")
stopifnot(all(names(train_set) %in% names(test)))
```

We can now align the training and the test sets

```{r}
test_set <- test[names(train_set)]
rm(test)
rm(essential, area, raion, demographic, macro)
```

### Transform variables

```{r}
source("../R/utilities.R")
train_tmp <- add_dummies(train_set[, -c(1, 2)], as_matrix = FALSE)
test_tmp <- add_dummies(test_set[, -c(1, 2)], as_matrix = FALSE)
train_set <- bind_cols(train_set[, 1:2], train_tmp)
test_set <- bind_cols(test_set[, 1:2], test_tmp)
rm(train_tmp, test_tmp)
cl_train <- map_chr(train_set, class)
cl_test <- map_chr(test_set, class)
stopifnot(all.equal(cl_train, cl_test))
```

The classes of all the columns match in the trianing set and in the test set. We save the training, prices and test sets as data frames.

```{r}
save(train_set, test_set, prices, 
     file = "../output/08-final_dataset.RData")
```

