# Further analysis of the round-priced transactions

We start adding the indicator variables for the 1M, 2M and 3M transactions. We want to understand whether there are raion-related patterns. For example, are this type of transactions more frequent in specific raions?

```{r}
library(magrittr)
library(tidyverse)
load("../output/08-final_dataset.RData")

prices %<>% mutate(
  round_price = as.numeric(price_doc %in% (c(1:3) * 1e6)),
  price_1m = price_doc == 1e06,
  price_2m = price_doc == 2e06,
  price_3m = price_doc == 3e06
)
```

I have checked whether the transactions with round prices take place always in the same raions, and it's not the case. I'm skipping this part of the code for brevity.

## Relative frequency of the round-priced transactions

```{r, rel_freq}
round_prices <- prices %>%
  group_by(year) %>%
  summarise(
    p_1m = sum(price_1m) / n(),
    p_2m = sum(price_2m) / n(),
    p_3m = sum(price_3m) / n()
  ) %>%
  gather(key = parameter, value = percentage, p_1m:p_3m)
```

Let's plot these results and see whether there's a pattern through time.

```{r}
p <- ggplot(
  data = round_prices,
  aes(x = year, y = percentage, fill = parameter)
  ) +
  geom_bar(stat = "identity", position = "dodge")
print(p)
```

## Subsampling strategy

We can, to be a bit conservative, take the average proportions for each type of round price from the last two years.

```{r last_two_years}
mean_14_16 <- round_prices %>% filter(year > 2013) %>%
  group_by(parameter) %>%
  summarise(mean_p = mean(percentage))
print(mean_14_16)
```

We now want to subsample the whole training set such that the round-priced transactions are present in these proportions. We don't modify the fraction of 3M prices because the estimated proportion is higher than the actual one.

```{r}
N <- nrow(train_set)

normal_prices <- filter(prices, round_price == 0)

prices_1m <- filter(prices, price_doc == 1e06)
prices_2m <- filter(prices, price_doc == 2e06)
prices_3m <- filter(prices, price_doc == 3e06)

sub_price_1m <- sample_n(prices_1m, round(N * mean_14_16$mean_p[1]))
sub_price_2m <- sample_n(prices_2m, round(N * mean_14_16$mean_p[2]))
# sub_price_1m <- sample_n(prices_3m, round(N * mean_14_16$mean_p[3]))
```

We now create a new training set containing these subsets

```{r}
prices <- bind_rows(
  normal_prices,
  sub_price_1m,
  sub_price_2m,
  prices_3m
)

sub_train_set <- train_set %>% filter(id %in% prices$id)
```

Let's verify that this is what we want

```{r}
prices %>%
  summarise(
    p_1m = sum(price_1m) / n(),
    p_2m = sum(price_2m) / n(),
    p_3m = sum(price_3m) / n()
  ) 
```

We save the subsampled training set in a new RData file.

```{r}
sub_train_set %<>% arrange(timestamp)
prices %<>% select(id:sub_area) %>% 
  arrange(timestamp)
```

## XGBoost model

```{r}
library(xgboost)
val_set <- sub_train_set %>% filter(timestamp >= "2015-02-01")
prices_val <- prices %>% filter(id %in% val_set$id)

sub_train_set %<>% filter(timestamp < "2015-02-01")
prices_train <- prices %>% filter(id %in% sub_train_set$id)

X_tr <- sub_train_set %>%
  mutate(year = lubridate::year(timestamp)) %>%
  select(-c(id, timestamp)) %>% as.matrix()
X_val <- val_set %>% 
  mutate(year = lubridate::year(timestamp)) %>%
  select(-c(id, timestamp)) %>% as.matrix()
X_test <- test_set %>%
  mutate(year = lubridate::year(timestamp)) %>%
  select(-c(id, timestamp)) %>% as.matrix()
X_train <- rbind(X_tr, X_val)

y_tr <- log(1 + prices_train$price_doc)
y_val <- log(1 + prices_val$price_doc)
y_train <- c(y_tr, y_val)

dTr <- xgb.DMatrix(X_tr, label = y_tr)
dVal <- xgb.DMatrix(X_val, label = y_val)
dTrain <- xgb.DMatrix(X_train, label = y_train)
dTe <- xgb.DMatrix(X_test)
```

Model fit 

```{r, results="hide"}
watchlist <- list(eval = dVal, train = dTr)

param <- list(
  eta = 0.1,
  nthread = 4,
  objective = "reg:linear",
  eval_metric = "rmse",
  gamma = 1,
  max_depth = 4,
  min_child_weight = 1,
  subsample = .8,
  colsample_by_tree = .8
)

tr <- xgb.train(
  params = param, 
  data = dTr, 
  nrounds = 300, 
  watchlist = watchlist
)

```

```{r}
pred_val <- predict(tr, dVal)
res_df <- tibble(y_val, pred_val)
p <- ggplot(res_df, aes(x = y_val, y = pred_val)) +
  geom_point(alpha = 0.1) + 
  geom_abline(slope = 1, intercept = 0) +
  geom_smooth()
print(p)
```

